### 2023-06-06 manuscripts/TCS2023/response.txt
Cleanup before addressing reviewer comments

We would first like to thank the reviewers for their insightful comments.

* Reviewer 1

"In my opinion, this submission is a solid paper, quite well written, and should
become a reference for anyone doing dynamics analysis of Boolean models. Below I
list several minor remarks, fixes, and suggestions for improvement. I believe
this work can be published after these minor corrections are implemented."

We fixed all the minor remarks highlighted by the reviewer, as can be seen in
red in the uploaded version.  Indeed it improved substantially the readability
of the article.