### 2023-06-06 manuscripts/TCS2023/response.txt
Done with reviewer 1's comments

We would first like to thank the reviewers for their insightful comments.

* Reviewer 1

"In my opinion, this submission is a solid paper, quite well written, and should
become a reference for anyone doing dynamics analysis of Boolean models. Below I
list several minor remarks, fixes, and suggestions for improvement. I believe
this work can be published after these minor corrections are implemented."

We thank the reviewer for their kind words and for a very thorough reading of
our article.  We fixed all the minor remarks highlighted by the reviewer, as can
be seen in red in the uploaded version.  We believe it improved substantially
the readability of the article.