### 2023-06-06 manuscripts/TCS2023/response.txt
Some answers for Reviewer 2

We would first like to thank the reviewers for their insightful comments.

* Reviewer 1

"In my opinion, this submission is a solid paper, quite well written, and should
become a reference for anyone doing dynamics analysis of Boolean models. Below I
list several minor remarks, fixes, and suggestions for improvement. I believe
this work can be published after these minor corrections are implemented."

We thank the reviewer for their kind words and for a very thorough reading of
our article.  We fixed all the minor remarks highlighted by the reviewer, as can
be seen in red in the uploaded version.  We believe it improved substantially
the readability of the article.

* Reviewer 2

"1. The authors employ an existing method of translating BNs to PNs. Anyway, they introduce a novel notion of a
specific class of siphons. It would be great to see more discussion on that general concept. Especially, I am not
familiar with the complexity of computing general siphons in PNs. It would be good to recall complexity issues
around that and to discuss (at least informally) the complexity of the problem of computing all conflict-free siphons
of the specific class of safe PNs."

TODO

"2. It is claimed that the PNs-based encoding (and also the results) are independent of the semantics of the update
scheme. It is not clear how this encoding could work with most-permissive semantics. I can imagine synchronous or
generally synchronous schemes to be embedded (still raising some questions on the complexity issues and
adaptability of the three methods presented in trappist)."

As explained in the article, and now extracted as a Corollary, the trap-spaces
of a Boolean network are indeed independent of any update scheme.  The
most-permissive semantics is not an update scheme (since it goes outside of the
Boolean state space) but an approximation technique for reachability and
trap-spaces.  As shown in the comparison with the mpbn tool that is based on
that semantics, the proposed method still outperforms any existing technique for
computing trap spaces, while also being able to handle non-locally-monotonic
networks.

"3. The paper is quite skeptic on the topic of using BDDs for identifying trap spaces. Anyway, BDDs have recently
proven to be useful for computing attractors (e.g., Benes et al. CAV 2021). I am aware the trap spaces are much more
complicated structures than (general) SCCs. Anyway, it might be good to see some deeper discussion that gives the
reader more information on why BDD would not work. Or the claim could be rather made weaker."

Since we do use BDDs in our implementation, in our case to obtain the Petri net
encoding, we are fully aware of their great characteristics.  As explained in
the introduction, what makes the current BDD-based approach, as implemented in
bioLQM, not efficient is that it tries to represent *all* trap spaces of a model
as a BDD and then to filter them to get minimal trap spaces. This is where the
inefficiency lies, and not in the specific use of BDDs per se.

We have added some points to highlight that we are not at any point claiming
that BDDs in general are not suited to this kind of task.

"4. At last of but not least, an investigative (may be naive) question comes to my mind: Are there (at the PN level) any
static analysis approaches that might help even (or improve) in identifying conflict-free siphons based on the
network structure? That would made the idea of the paper even more promising."

We are not aware of any such technique, but have now highlighted (at the end of
Section 3 and in conclusion) that the link between Petri nets and Trap spaces
could bring novelties both ways.

"The critical problem I see in this submission is that several parts of the related work are missing (especially the
coverage of TCS background is quite low).
First, I think there must be definitely more background on the related problems in the domain of PNs. There exist
several approaches to compute siphons. I am not from the PN domain but by some fast googling I see many papers
on computing siphons. It would be good to include some of them in the background and discuss if they might be also
considered for computing conflict-free siphons. I can believe the simple argument is that existing approaches are
general and hence cannot be that efficient as the approach presented in the paper. Anyway, that should be
discussed explicitly to add some more rigour."

As explained at the very end of Section 3, there is, to our knowledge, no method
to compute maximal generic siphons, and even less maximal conflict-free
siphons. We refer to a recent survey [34] but as it shows, the focus is on the
minimal siphons, and of course generic ones.  We have added a sentence to
explain even more why these methods are not adapted to our problem.

"Second, in my opinion some work on attractor identification might be seemed as strongly related since there are
several useful concepts that have strong relation to trap spaces. In particular, there is the quite recent phd work of
Fontanels et al. elaborating on the concept of phenotype control where I think that phenotypes can somehow gain
from computing the trap spaces. Also the network reductions as recently introduced by Tonello et al. (arxiv 2023)
seem to be promising for shifting the scalability further. This comes to question especially on lines 692-698."

TODO

We would like to thank the reviewer for their useful comments. We also addressed all the minor issues they pointed out.

Minor issues
==========
page 5. The general definition of BNs semantics at the line 35 is quite non-standard. I understand the attempt to
capture a sort of subclass of update modes abstractly but I think it still would make sense to be more explicit and
give some concrete instances (that can be discussed throughout the paper). In the current version I have the feeling
you focus primarily on asynchronous update mode.

Do we have a reference for this?

Figure 1: add the topology of the BN

TODO

line 172: there is the abuse of notion $m$ (already used for "markings" of BNs)

TODO
m = subspaces = markings is indeed not nice